---
title: "Movie recommendation"
author: "Shobhit Tiwari"
output:
  html_document: default
  html_notebook: default
data: January 21, 2018
---


```{r include=FALSE}
library(stringr)
library(reshape2)
library(recommenderlab)
```

## Synopsis

In this project we will predict missing ratings for all users and also recommend 10 movies for each user.We will use MovieLense dataset that for building our recommender.

Lets quickly load and explore the dataset

```{r}

data(MovieLense)
movMat<- as(MovieLense, "matrix")
dim(movMat)

```


So we have a rating matrix of 943 users and 1663 movies. Lets check no of available ratings in the matrix


```{r}
prod(dim(movMat)) - sum(is.na(movMat))

```


So we can see around 100k ratings are present in the dataset. 


Let's see the ratings by user 1 

```{r}
head(movMat[1,])

```

Lets quickly see the histogram of ratings per user and movie 


```{r}
## number of ratings per user
hist(rowCounts(MovieLense), main = "Ratings per movie")

## number of ratings per movie
hist(colCounts(MovieLense), main = "Ratings per user")

```

We can see that very few movies are having high no of ratings and very few users have rated a large no of movies.

Now let us see a heatmap visualization of part of the  dataset

```{r}
## visualize part of the matrix
image(MovieLense[1:100,1:100])


```


Now we will split data into trian and test set using evaluation scheme function in the package.

 
```{r}
set.seed(1233)
e <- evaluationScheme(MovieLense, method="split", train=0.9, given=-5, goodRating=4)

```


We are splitting data into two parts training containing 90% of the users and 10% users are kept for test set with 5 ratings as unknown and rest as known. 


First we will build a model using user based collaborative filtering. In UBCF rating for a user is predicted by finding similarity between user ratins and then taking weighted average of user raitngs to predict the missing ratings for the user. 

Having users who give high (or low) ratings to all their movies might bias the results.
We will normalize the ratings to minimize user bias.

```{r}
#Create UBCF Recommender Model. UBCF stands for User-Based Collaborative Filtering
UBCF_model <- Recommender(getData(e, "train"), 
                                 method = "UBCF", 
                                 param=list(normalize = "Z-score",
                                                 method="Cosine",
                                                 nn=50))
```

Now we will see top 10 movie recommendation for user 1 


```{r}

firstUserPred <- predict(UBCF_model, 
                 MovieLense[1,], 
                 n=10, type = "topNList") 
firstUserPred <- as(firstUserPred, "list")
# Get recommendation for first user only top 10 movies 
firstUserPred[[1]]

```


Now predicting for the test set and checking accuracy of model

```{r}
recom <- predict(UBCF_model, 
                 getData(e, "known"), 
                 n=10, type = "ratings") 
calcPredictionAccuracy(recom, getData(e, "unknown"))
```


Secondly we will build an item based model. In this approach we predict ratings for a movie finding similar movies and taking weighted average of their ratings. 


```{r}
IBCF_model <- Recommender(getData(e, "train"), 
                                 method = "IBCF", 
                                 param=list(normalize = "Z-score",
                                                 method="Cosine",
                                                 k=50))

```


Now we will see top 10 movie recommendation for user 1 


```{r}

firstUserPred.IBCF <- predict(IBCF_model, 
                 MovieLense[1,], 
                 n=10, type = "topNList") 
firstUserPred.IBCF <- as(firstUserPred.IBCF, "list")
# Get recommendation for first user only top 10 movies 
firstUserPred.IBCF[[1]]

```



Lets check on test set and calculate the RMSE 


```{r}

recom.IBCF <- predict(IBCF_model, 
                 getData(e, "known"), 
                 n=10, type = "ratings") 
calcPredictionAccuracy(recom.IBCF, getData(e, "unknown"))

```


Lets build a matrix factorization model for recommendation using Alternating least squares model

```{r}

als_model <- Recommender(getData(e, "train"), 
                                 method = "ALS", 
                                 param=list(normalize = "Z-score",lambda=0.1, n_factors=200, 
                                  n_iterations=10, seed = 214, verbose = FALSE))


```


Now predicting for the test set and checking accuracy of model

```{r}
recom_als <- predict(als_model, 
                 getData(e, "known"), 
                 n=10, type = "ratings") 
calcPredictionAccuracy(recom_als, getData(e, "unknown"))
```


Now we will see top 10 movie recommendation for user 1 

```{r}
firstUserals_Pred <- predict(als_model, newdata =  MovieLense[1,],n=10, type = "topNList") 
firstUserals_Pred <- as(firstUserals_Pred, "list")
# Get recommendation for first user only top 10 movies 
firstUserals_Pred[[1]]

```




## Conclusion

As we can see from our analysis user based collaborative filterting gives the best result in this case. 